---
title: "Stats2 Project1"
author: "Andrew Leppla"
date: "2/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import and Inspect Beer Reviews Data
```{r import and inspect}
beer_reviews = read.csv('~/MDS-6XXX-Stats2/2020-Stats2-Project-Beer-Reviews/beer_reviews.csv')
#head(beer_reviews)
#str(beer_reviews)

```

## Recode review_time variable for Time Series

Per the website summary, "This dataset consists of beer reviews from Beeradvocate. The data span a period of more than 10 years, including all ~1.5 million reviews up to November 2011." https://data.world/socialmediadata/beeradvocate

Assuming the review_time is in seconds, the reviews actually span over 15 years.  A likely date for t=0 is Jan 1, 1970 which puts the last review at Jan 11, 2012 rather than November 2011.  In the absence of any other information, this is close enough to do time series analysis.  

The data are not equally spaced over time which is not surprising.  People aren't reviewing beers every second of every day in a predominantly U.S.-based market.   

To address the unequal data spacing over time, reviews were averaged by Month as well as by Week. 

```{r recode review_time}

library(dplyr)
library(tidyverse)
library(ggplot2)

#Investigate review_time for recoding
max_time = max(beer_reviews$review_time) #November 2011?
min_time = min(beer_reviews$review_time) 
diff = (max_time- min_time)/60/60/24/365.25 #Over 10 years?
diff #Actually over 15 years

#Recode review_time with the common origin time Jan 1, 1970
beer_reviews$review_date_time = as.POSIXct.numeric(beer_reviews$review_time,origin='1970-01-01 00:00:00',tz='EST')
head(beer_reviews$review_date_time)
max(beer_reviews$review_date_time) #Jan 11, 2012
min(beer_reviews$review_date_time) #Aug 21, 1996

#############################################################################################
#Explore Overall reviews vs. time for 90 Minute IPA (most-rated beer)
beer_reviews %>% filter(review_date_time>'2011-01-01 00:00:00' & beer_name=='90 Minute IPA') %>% ggplot(aes(x=review_date_time,y=review_overall,color=beer_name)) + geom_line()
#Data is not equally spaced over time (as expected)
#############################################################################################

#Average by month (and by week) for equally spaced data over time
beer_reviews$time_vectors = as.POSIXlt.numeric(beer_reviews$review_time,origin='1970-01-01 00:00:00',tz='EST') #Splits the time "YYYY-MM-DD HH:MM:SS" into vectors "YYYY","MM","DD",etc.
beer_reviews$Year = as.numeric(beer_reviews$time_vectors$year) + 1900 #Calling $year from class POSIXlt resets the origin to year 1900 (see ?POSIXlt)
beer_reviews$Mon = as.numeric(beer_reviews$time_vectors$mon) + 1 # $mon from POSIXlt is from 0-11, add 1 so it's from 1-12
beer_reviews$Week = as.integer((beer_reviews$time_vectors$yday + 1)/7 +1) #Divide days (1-366, 366=leap day) by 7 and round to nearest integer for weeks 1-53 (not perfect, close enough)

```

## Monthly Time Series - 90 Minute IPA

Dogfish Head's "90 Minute IPA" has the most reviews of any beer in the data set.  This was used for Time Series Analysis.  

The data look fairly stationary, but there is a possible upward trend for the last 20 months.     

The monthly data are highly variable for the first 5-10 months. This is because the reviews are too few (n= 1-7) to get a reliable average.  

```{r Monthly Time Series - 90 Min IPA}
library(tseries)
library(forecast)

#Need to recode the time variable back to seconds to pipe it through group_by() and summarize()
beer_reviews$time_vectors = as.POSIXct.numeric(beer_reviews$review_time,origin='1970-01-01 00:00:00',tz='EST') 

#Filter for 90 Minute IPA Average Monthly Reviews
tseries_90minIPA_Mon = beer_reviews %>% filter(beer_name=='90 Minute IPA') %>% group_by(Year, Mon) %>% summarize(Overall=mean(review_overall),Count=n()) #class=tibble
tseries_90min_Mon_df=data.frame(tseries_90minIPA_Mon) #convert tibble to data frame

#Plot the data before modeling
##Data has nonconstant variance for months 1-7 with low n
tseries_90min_Mon_df$Index=1:nrow(tseries_90min_Mon_df)
tseries_90min_Mon_df %>% ggplot(aes(x=Index,y=Overall)) + geom_line() + geom_point() + ylab("Mean Overall Rating") + xlab("Months from Dec. 2001")

#ACF and PACF Plots
par(mfrow=c(1,2))
Acf(tseries_90min_Mon_df$Overall,lag.max=30) #Up to lag 19 looks significant
Pacf(tseries_90min_Mon_df$Overall,lag.max=30) #Up to lag 19 looks significant

par(mfrow=c(1,1))
```


## Monthly Time Series - Filter out the first 7 months for constant variance

Manual model AR_1_MA_2_13 is the best fit with the lowest AIC and RMSE. 

```{r Time Series Analysis - 90 Min IPA}
#Plot that filters out the first 7 months with
#nonconstant variance and insufficient sample sizes
tseries_90min_Mon_df %>% filter(Index>7) %>% ggplot(aes(x=Index,y=Overall)) + geom_line() + geom_point() + ylab("Mean Overall Rating") + xlab("Months from August 2002") 

IPA90min_Mon_filtered = tseries_90min_Mon_df %>% filter(Index>7) 
min(IPA90min_Mon_filtered$Year) #August 2002

par(mfrow=c(1,2))
Acf(IPA90min_Mon_filtered$Overall,lag.max=35) #Has a longer lag = 13 
Pacf(IPA90min_Mon_filtered$Overall,lag.max=35)#Has a longer lag = 13 

par(mfrow=c(1,1))

#Auto ARIMA

ARIMA.1<-auto.arima(IPA90min_Mon_filtered$Overall)
summary(ARIMA.1) #MA(1), missing lag 13, AIC = -130.2
plot(forecast(ARIMA.1,h=10))
points(1:length(IPA90min_Mon_filtered$Overall),fitted(ARIMA.1),type="l",col="blue")
tsdisplay(residuals(ARIMA.1),lag.max=20,main="MA(1) Resid. Diagnostics")
tsdisplay(residuals(ARIMA.1),lag.max=115,main="MA(1) Resid. Diagnostics")

ARIMA.2<-auto.arima(IPA90min_Mon_filtered$Overall,stepwise=F)
summary(ARIMA.2) #AR(2), missing lag 13, AIC = -131.2
plot(forecast(ARIMA.2,h=10))
points(1:length(IPA90min_Mon_filtered$Overall),fitted(ARIMA.2),type="l",col="blue")
tsdisplay(residuals(ARIMA.2),lag.max=20,main="AR(2) Resid. Diagnostics")

#Manual ARIMA

MA_13 = arima(IPA90min_Mon_filtered$Overall, order=c(0,0,13))
summary(MA_13) #AIC = -134.6
plot(forecast(MA_13,h=20))
points(1:length(IPA90min_Mon_filtered$Overall),fitted(MA_13),type="l",col="blue")
tsdisplay(residuals(MA_13),lag.max=110,main="MA(13) Resid. Diagnostics")

AR_13 = arima(IPA90min_Mon_filtered$Overall, order=c(13,0,0))
summary(AR_13) #AIC = -131.9
plot(forecast(AR_13,h=70))
points(1:length(IPA90min_Mon_filtered$Overall),fitted(AR_13),type="l",col="blue")
tsdisplay(residuals(AR_13),lag.max=110,main="AR(13) Resid. Diagnostics")

MA_1_13 = arima(IPA90min_Mon_filtered$Overall, order=c(0,0,13), fixed=c(NA,0,0,0,0,0,0,0,0,0,0,0,NA,NA))
summary(MA_1_13) #AIC = -150.5
plot(forecast(MA_1_13,h=20))
points(1:length(IPA90min_Mon_filtered$Overall),fitted(MA_1_13),type="l",col="blue")
tsdisplay(residuals(MA_1_13),lag.max=110,main="MA(1,13) Resid. Diagnostics")

AR_2_13 = arima(IPA90min_Mon_filtered$Overall, order=c(13,0,0),fixed=c(NA,NA,0,0,0,0,0,0,0,0,0,0,NA,NA))
summary(AR_2_13)
plot(forecast(AR_2_13,h=50))
points(1:length(IPA90min_Mon_filtered$Overall),fitted(AR_2_13),type="l",col="blue")
tsdisplay(residuals(AR_2_13),lag.max=110,main="MA(1,13) Resid. Diagnostics")

#Model with best AIC and RMSE
AR_1_MA_2_13 = arima(IPA90min_Mon_filtered$Overall, order=c(13,0,13),fixed=c(NA,0,0,0,0,0,0,0,0,0,0,0,0,NA,NA,0,0,0,0,0,0,0,0,0,0,NA,NA))
summary(AR_1_MA_2_13)
plot(forecast(AR_1_MA_2_13,h=30))
points(1:length(IPA90min_Mon_filtered$Overall),fitted(AR_1_MA_2_13),type="l",col="blue")
tsdisplay(residuals(AR_1_MA_2_13),lag.max=110,main="AR(1)MA(2,13) Resid. Diagnostics")
```

##Differencing 

These models don't capture the significant lag 13 residuals and may be overfitting the data.

```{r}

#Differencing degree of 1 (d=1)
#Stationary model per the forecast plot
ARIMA.d1<-auto.arima(IPA90min_Mon_filtered$Overall,d=1)
summary(ARIMA.d1)  
plot(forecast(ARIMA.d1,h=15))
points(1:length(IPA90min_Mon_filtered$Overall),fitted(ARIMA.d1),type="l",col="blue")
tsdisplay(residuals(ARIMA.d1),lag.max=20,main="MA(1) Resid. Diagnostics")

#Differencing degree of 2 (d=2)
#Nonstationary model per the forecast plot
ARIMA.d2<-auto.arima(IPA90min_Mon_filtered$Overall,d=2,stepwise=F)
summary(ARIMA.d2)  
plot(forecast(ARIMA.d2,h=15))
points(1:length(IPA90min_Mon_filtered$Overall),fitted(ARIMA.d2),type="l",col="blue")
tsdisplay(residuals(ARIMA.d1),lag.max=20,main="MA(1) Resid. Diagnostics")

#Manual Differencing
##This appears to be doing 13 degrees of differencing rather than differencing for lag 13 

diff.data<-arima(IPA90min_Mon_filtered$Overall,order=c(0,13,0))
summary(diff.data)
tsdisplay(residuals(diff.data),lag.max=30,main="Resid. Diagnostics 1st Order Difference")

Diff1 = auto.arima(diff.data$residuals)
summary(Diff1) #AR(3)
plot(forecast(Diff1,h=10))
points(1:length(diff.data$residuals),fitted(Diff1),type="l",col="blue")
tsdisplay(residuals(Diff1),lag.max=105,main="AR(3) Resid. Diagnostics")

Diff2 = auto.arima(diff.data$residuals,stepwise=F)
summary(Diff2) #AR(1)MA(4)
plot(forecast(Diff2,h=10))
points(1:length(diff.data$residuals),fitted(Diff2),type="l",col="blue")
tsdisplay(residuals(Diff2),lag.max=105,main="AR(1)MA(4) Resid. Diagnostics")
```



#The remaining analysis in this Rmd file is questionable due to the nonconstant variance

## Auto ARIMA Model Selection - Nonconstant Variance with months 1-7

Based on the AIC selection criteria, the AR(4)MA(2) model is the best fit that keeps the residuals at or below the significance threshold.

The residual plots of ACF and PACF don't decay until after lag 50 and 60, respectively.  
-Did IPAs become more popular/trendy around this time?
-Was there a surge of reviews?  
-Was there a change in the product's publicity or a new marketing campaign?  

```{r}
par(mfrow=c(1,1))

#Auto ARIMA fits an AR(2)MA(2) model
#but misses the significant PACF residual at lag 19 
ARIMA.fit<-auto.arima(tseries_90min_Mon_df$Overall)
ARIMA.fit #AR(2)MA(2) model, AIC=-83.3
summary(ARIMA.fit)
plot(forecast(ARIMA.fit,h=20))
points(1:length(tseries_90min_Mon_df$Overall),fitted(ARIMA.fit),type="l",col="blue")

#Lag 19 residual in PACF
tsdisplay(residuals(ARIMA.fit),lag.max=20,main="AR(2)MA(2) Resid. Diagnostics")

#Residuals decay after lag 50-60.
#Some evidence of non-stationary behavior between the first and second halves of the data?
tsdisplay(residuals(ARIMA.fit),lag.max=120,main="AR(2)MA(2) Resid. Diagnostics")


#Increase the allowable max order and remove stepwise selection criteria to get higher order models
ARIMA.fit1<-auto.arima(tseries_90min_Mon_df$Overall,max.order=70,stepwise=F,ic=c('aic'))
summary(ARIMA.fit1) #AR(4)MA(2) model, AIC=-86.5, BIC=-64.2
plot(forecast(ARIMA.fit1,h=20))
points(1:length(tseries_90min_Mon_df$Overall),fitted(ARIMA.fit1),type="l",col="blue")

#Lag 15 and 19 are now just under/at the significance threshold in ACF & PACF
tsdisplay(residuals(ARIMA.fit1),lag.max=20,main="AR(4)MA(2) Resid. Diagnostics")

#Residuals still decay after lag 50-60.
tsdisplay(residuals(ARIMA.fit1),lag.max=120,main="AR(4)MA(2) Resid. Diagnostics")


#Try BIC for non-stepwise selection criteria
ARIMA.fit1b<-auto.arima(tseries_90min_Mon_df$Overall,max.order=20,stepwise=F,ic=c('bic'))
summary(ARIMA.fit1b) #AR(2) model, AIC=-84.6, BIC=-73.5
plot(forecast(ARIMA.fit1b,h=20))
points(1:length(tseries_90min_Mon_df$Overall),fitted(ARIMA.fit1b),type="l",col="blue")

#Lag 19 is significant in Residual PACF
tsdisplay(residuals(ARIMA.fit1b),lag.max=20,main="AR(2) Resid. Diagnostics")


ARIMA.log<-auto.arima(log(tseries_90min_Mon_df$Overall))
summary(ARIMA.log)
plot(forecast(ARIMA.log,h=20))
points(1:length(tseries_90min_Mon_df$Overall),fitted(ARIMA.log),type="l",col="blue")
```

## Try to manually tune the model for data with nonconstant variance

Improved the residual lag plots somewhat at the expense of AIC.

ARIMA.fit3 and ARIMA.fit4 appear to be the best manual models.

```{r}
#Manually try to improve the model & residual decay 

ARIMA.fit2=arima(tseries_90min_Mon_df$Overall, order=c(17,0,2))
summary(ARIMA.fit2) #AIC = -77.0
plot(forecast(ARIMA.fit2,h=12))
points(1:length(tseries_90min_Mon_df$Overall),fitted(ARIMA.fit2),type="l",col="blue")
tsdisplay(residuals(ARIMA.fit2),lag.max=120,main="ARMA Resid. Diagnostics")

ARIMA.fit3=arima(tseries_90min_Mon_df$Overall, order=c(0,0,17))
summary(ARIMA.fit3) #AIC = -82.0
plot(forecast(ARIMA.fit3,h=12))
points(1:length(tseries_90min_Mon_df$Overall),fitted(ARIMA.fit3),type="l",col="blue")
tsdisplay(residuals(ARIMA.fit3),lag.max=120,main="MA(17) Resid. Diagnostics")

ARIMA.fit4=arima(tseries_90min_Mon_df$Overall, order=c(2,0,15))
summary(ARIMA.fit4) #AIC=-84.6 
par(mfrow=c(1,1))
plot(forecast(ARIMA.fit4,h=12))
points(1:length(tseries_90min_Mon_df$Overall),fitted(ARIMA.fit4),type="l",col="blue")
tsdisplay(residuals(ARIMA.fit4),lag.max=120,main="AR(2)MA(15) Resid. Diagnostics")

ARIMA.fit5=arima(tseries_90min_Mon_df$Overall, order=c(4,0,11))
summary(ARIMA.fit5) #AIC=-85.1, lag 7 is close to the sig. line
plot(forecast(ARIMA.fit5,h=30))
points(1:length(tseries_90min_Mon_df$Overall),fitted(ARIMA.fit5),type="l",col="blue")
tsdisplay(residuals(ARIMA.fit5),lag.max=120,main=" Resid. Diagnostics")

```


## Model non-stationary behavior ?

This didn't add any value to the model.

```{r}
OLS_Year<-arima(tseries_90min_Mon_df$Overall,order=c(0,0,0),xreg=tseries_90min_Mon_df$Year)
summary(OLS_Year)
tsdisplay(residuals(OLS_Year),lag.max=120,main="Resid. Diagnostics of OLS")

ARIMA_Year=auto.arima(tseries_90min_Mon_df$Overall, xreg=beer_tseries_df$Year,max.order=10,stepwise=F)
summary(ARIMA_Year)
tsdisplay(residuals(ARIMA_Year),lag.max=120,main="Year + AR(4)MA(2) Resid. Diagnostics") #Still have same model & issue, lag 19 residuals

ARMA_Year1<-arima(tseries_90min_Mon_df$Overall,order=c(2,0,11),xreg=tseries_90min_Mon_df$Year)
summary(ARMA_Year1)
tsdisplay(residuals(ARMA_Year1),lag.max=120,main="Resid. Diagnostics of OLS")


#Seasonal differencing (D=1) - doesn't appear to do anything
ARIMA_delta=auto.arima(tseries_90min_Mon_df$Overall, D=1,max.order=15,stepwise=F) 
ARIMA_delta
tsdisplay(residuals(ARIMA_delta),lag.max=120,main=" Resid. Diagnostics")

```


## 90 Minute IPA by Week



```{r}
tseries_90minIPA_Wk = beer_reviews %>% filter(beer_name=='90 Minute IPA') %>% group_by(Year, Week) %>% summarize(Overall=mean(review_overall),Count=n())
tseries_90min_Wk_df=data.frame(tseries_90minIPA_Wk)

tseries_90min_Wk_df$Index=1:nrow(tseries_90min_Wk_df)
tseries_90min_Wk_df %>% filter(Index>1000) %>% ggplot(aes(x=Index,y=Overall)) + geom_line() + ylab("Mean Overall Rating")

par(mfrow=c(1,2))
Acf(tseries_90min_Wk_df$Overall,lag.max=120) 
Pacf(tseries_90min_Wk_df$Overall,lag.max=120) 

par(mfrow=c(1,1))


ARIMA.Wk<-auto.arima(tseries_90min_Wk_df$Overall)
summary(ARIMA.Wk)
plot(forecast(ARIMA.Wk,h=10))
points(1:length(tseries_90min_Wk_df$Overall),fitted(ARIMA.Wk),type="l",col="blue")

#Residual lag 23 is significant
tsdisplay(residuals(ARIMA.Wk),lag.max=30,main="ARIMA(4,1,1) Resid. Diagnostics")

#Lots of hits for lag < 1000, decays eventually
tsdisplay(residuals(ARIMA.Wk),lag.max=2000,main="ARIMA(4,1,1) Resid. Diagnostics")


#Turn off stepwise selection for more model terms
ARIMA.Wk1<-auto.arima(tseries_90min_Wk_df$Overall,stepwise=F)
summary(ARIMA.Wk1)
plot(forecast(ARIMA.Wk1,h=10))
points(1:length(tseries_90min_Wk_df$Overall),fitted(ARIMA.Wk1),type="l",col="blue")

#Residual lag 23 is still significant
tsdisplay(residuals(ARIMA.Wk1),lag.max=30,main="ARIMA(4,1,1) Resid. Diagnostics")

tsdisplay(residuals(ARIMA.Wk1),lag.max=2000,main="ARIMA(4,1,1) Resid. Diagnostics")

```

## 90 Minute IPA by Index

Data is too noisy and discrete to model

```{r}
IPA90Min = beer_reviews %>% filter(beer_name=="90 Minute IPA")
IPA90Min$Index=1:nrow(IPA90Min)

#First review of 90 Minute IPA is in 2001
as.POSIXct(min(IPA90Min$review_time),origin='1970-01-01 00:00:00',tz='EST')

IPA90Min %>% filter(Index>3000) %>% ggplot(aes(x=Index,y=review_overall)) + geom_line() + geom_point() + ylab("Overall Rating")

par(mfrow=c(1,2))
Acf(IPA90Min$review_overall,lag.max=120) #Decays after
Pacf(IPA90Min$review_overall,lag.max=120) 

par(mfrow=c(1,1))

ARIMA.Index<-auto.arima(IPA90Min$review_overall)
summary(ARIMA.Index)
plot(forecast(ARIMA.Index,h=50))
points(1:length(IPA90Min$review_overall),fitted(ARIMA.Index),type="l",col="blue")

#Significant ACF residuals at lags 2-6  
tsdisplay(residuals(ARIMA.Index),lag.max=25,main="Resid. Diagnostics")
tsdisplay(residuals(ARIMA.Index),lag.max=3200,main="Resid. Diagnostics")

```
